### 2023-05-02 Expérimentations (Théa)/2. Expérimentations (Bookdown)/Bookdown/03-theoritical-frame.Rmd

---
knit: bookdown::render_book
editor_options: 
  markdown: 
    wrap: 80
---

# Theoretical framework

Family as a unit of analysis can be particularly fruitful when studying health
inequalities in the context of a societal-wide crisis. Through the lens of the
COVID-19 pandemic, the following points illustrate why exploring family dynamics
provides additional insight into the underlying mechanisms of health
disparities. First, the family is the most fundamental unit of social
organization. It facilitates access to resources and the transmission of social
and psychological capital from one generation to the next. Therefore, a better
analysis of how crises and adverse life events affect people requires a better
understanding of how families relate to each other and interact with external
forces. Second, during public health emergencies, families may be thrown into
different structures due to unprecedented financial, mental, and physical
challenges. The risk of family dissolution, for example, can be heightened when
members must spend unusually long periods together at home and have limited
opportunities to engage with friends and the wider community. Conversely,
families can develop closer bonds through self-sacrifice, shared
responsibilities, and increased care and support. As the dependency status of
seniors increases, the elderlies are also more likely to be integrated into
formerly nucleus families. We posit that the extent to which a family can adapt
its functions to meet new challenges is a predictor of the well-being of its
members and the resilience of its broader population. Third, the family is a
crucial link to social institutions, such as the health care system, schools,
and government agencies whose decisions have significant downstream impacts.
Families are viewed by many public-level entities as a reference when allocating
resources, enforcing new standards, and responding to emerging needs and
demands. From this perspective, the family operates at the nexus of individual
and societal structures, embedding the tripple effects of public policy in its
most consequential ways. Finally, the family has a crucial role to play in
determining the life course of its members. As COVID-19 elevates the importance
of familial bonds, the post-pandemic period might be marked by an increased
influence of families on the lives of their members and the social structures
they inhabit.

The above conceptualization echoes several other approaches to studying
long-term health inequalities. For instance, life course theory emphasizes the
significance of family ties and social environment as they structure prospects
for individuals across their life span (Elder, 1977; Elder Jr., 1998; Macmillan,
2005; Settersten et al., 2020). Likewise, social capital theory and the
stress-buffering hypothesis consider family systems a crucial factor in health
and well-being (Choi and Jun, 2020; Cohen and McKay, 1984; Cohen and Wills,
1985; Harpham et al., 2004; Kawachi and Berkman, 2001; Lavee et al., 1987;
Mandelbaum et al., 2020). They show that the family can offer tangible (e.g.,
food, shelter, and money) and intangible resources (e.g., support, social
status, psychological comfort, and institutional advantages), protecting its
members from adverse life events. Similarly, the social-ecological model and the
theory of fundamental causes highlight how the SES, familial and extrafamilial
factors pattern people's life trajectories and health outcomes (Barkan and
Rocque, 2018; Clouston and Link, 2021; Figge et al., 2018; Kim, 2020; Libman et
al., 2012; Ungar, 2015). However, all these approaches are ostensibly focused on
life periods where the general social context is relatively stable. In this
article, we present our conceptualization of the family in times of
unprecedented social upheaval. Specifically, our framework seeks to characterize
family dynamics as a suitable lens to analyze health inequalities in the face of
a virulent public health crisis.

\newpage