### 2023-05-02 Expérimentations (Théa)/2. Expérimentations (Bookdown)/Bookdown/05-results.Rmd

---
knit: bookdown::render_book
editor_options: 
  markdown: 
    wrap: 80
---

# 3. Results

## 3.1. Sample distribution

## 3.2. Exposure by immigration status

Figure 1 presents the exposure to the pandemy accroding to the immigrants and
minority visible status. For all the six measures of exposure, it appears that
the majority group is the most exposed to the COVID-19 in all instance. The big
difference between the groups is about the exposure trhough working outside.
Around 50% of the member of this group or a family member work outside the home
at the time of the survey. This percentage is only 30% for the racialised
immigrants and 25% for the non-racialised immigrants.

![Figure 1.1: Exposure](Figures/Exposition1.png){width="60%"}

![Figure 1.2: Explicative model](Figures/Exposure.jpeg)

## 3.3. Consequences by immigrations status

The Figure 2 presents the consequences of the COVID-19 for a family member to
access to ressources. It shows that in all accounts but one (getting medecine),
the racialised immigrants express more diffciult to assess ressources. Around
20% of them stated having difficulty to get health care compare, around 10% for
obtaining food and 7% for obtaining family insurrance. Comparatively, these
proportion are respectively 13%, 7% and 4% for people and thier family member in
the majority group and 19%, 3% and 4% respectively among the racialised
immigrants.

![Figure 3.1: Consequences on accessing
ressources](Figures/Famille1.png){width="80%"}

![Figure 3.2: Explicative model](Figures/Ressources.jpeg)

## 3.4. Consequences on performing usual duties

The Figure 4 compares the response from the three groups on performing usual
duties. It appears that for all groups the situation caused by the COVID-19 has
inexpectedly improved the situation in various way, irrespective of the
immigrants status of the respondents.

![Figure 4.1: Consequences on performing usual duties](Figures/Entente1.png)

![Figure 4.2: Explicative model](Figures/Abilities.jpeg)

## 3.5. Consequences on employment by immigration status

The consequences on employment are mixed, with the severe consequences more
prevalent among the racialised immigrants. The Figure 3 shows for instance that
close to 45% of the racialised immigrants have stated that their family income
has decreased compare to just 33% of the majority group and 37% of the
non-racialed immigrants. In addition, 11% of the racialised immigrants reveals
having loss their job compared to 9% among the non-racialised immigrants and 2%
among members of the majority group. At the opposite, the members of the
majority groups are more likely to report a reduction of the working hours or
for stopping momentarily their work.

![Figure 5.1: Consequences on employment and income](Figures/Emploi1.png)

![Figure 5.2: Explicative model](Figures/Employment.jpeg)

## 3.7 Help received by immigrations status

Finally, we compared the different group on benefitting of the government
assistance. It appears that racialised immigrants with 35% of the case have
expressed having received the Canada Emergency Response Benefit wheras this
proportion is 31% for the non-racialised immigrants and 25% for the member of
the majority group. For the province Temporary Aid for Workers Program, around
5% of the respondents have received it.

![](Figures/Aide1.png)

![Figure 3.2: Explicative model](Figures/Government.jpeg)

\newpage