### 2023-05-24 Codes/Sentiment_an1.Rmd

---
title: "Sentiment_an1"
author: "Thea"
date: "2023-05-24"
output: html_document
---