### 2023-05-31 2. Trajipaq_covid/2. Codes/Trajipaq_covid_reg2.Rmd

---
title: "Familial consequences of the Covid-19 in Québec: Comparison between immigrants and natives"
subtitle: "Conséquences familiales de la Covid-19 au Québec: comparaison immigrants - natifs"
author:
- Vissého Adjiwanou ^[corresponding author - adjiwanou.vissého@uqam.ca]
- Département de Sociologie, Université du Québec à Montréal (UQAM)
- Solène Lardoux (Département de démographie, Université de Montréal)
- Nima Zahedinameghi (Département de Sociologie, UQAM)
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document: 
    toc: true
    toc_depth: 3
  word_document: default
  template: svm-latex-ms2.tex 
  pdf_document: default
  citation_package: natbib
bibliography: Covid_Canada.bib
---

<!--
Ce fichier est pour les modèles de régression
Reg2
Modèle avec la variable imm_minorite
-->

```{r, echo=FALSE, warning=FALSE}

rm(list = ls())

library(tidyverse)
library(summarytools)
library(srvyr)
library(survey)
library(ggpubr)
library(writexl)
library(stargazer)
library(modelsummary)
library(labelled)

trajipaq_covid <- read_rds("../Data/trajipaq_covid.RData")

labeled_vars <- sapply(trajipaq_covid, function(x) class(x) == "labelled")
trajipaq_covid[, labeled_vars] <- lapply(trajipaq_covid[, labeled_vars], as_factor)

# Base de données excluant others
trajipaq_covid <-
  trajipaq_covid %>% 
  filter(imm_minorite != "Others")

freq(trajipaq_covid$imm_minorite)
freq(trajipaq_covid$statut_mat)

freq(trajipaq_covid$length_stay_2)

```

## Recodage des variables

## Pondération

```{r}

#trajipaq_nm <- na.omit(trajipaq_covid)
#trajipaq.design <- svydesign(id = ~ CaseId, weights = ~ Poidstot1, data = trajipaq_nm)
trajipaq.design <- svydesign(id = ~ CaseId, weights = ~ Poidstot1, data = trajipaq_covid)

```

## Modèle de régression logistique

### Les différents modèles

```{r}

#expo <-
#  trajipaq_covid %>% 
#  dplyr::select(e_hors_dom, e_four_soin, e_quarantaine, e_expose, e_symptome, e_hospitalise, e_intensif, e_decede)

#dfSummary(expo, 
#          plain.ascii  = FALSE, 
#          style        = "grid", 
#          graph.magnif = 0.75, 
#          valid.col    = FALSE,
#          tmp.img.dir  = "/tmp")

freq(trajipaq_covid$statut_imm)
freq(trajipaq_covid$imm_minorite)

dep1 <- list("e_hors_dom~", "e_four_soin~", "e_quarantaine~", "e_expose~",
            "e_symptome~", "e_hospitalise~", "e_intensif~", "e_decede~")

dep2 <- list("expo~")

indep1 <- list("relevel(imm_minorite, 3)")

indep2 <- list("relevel(imm_minorite, 3) + relevel(sexe, 2) + length_stay_2 + gpage + niv_etude + statut_mat + statut_emploi")

indep2_bis <- list("relevel(imm_minorite, 3) + relevel(sexe, 2) + length_stay_2 + gpage + niv_etude + statut_mat")   # For emploi

# Change of the reference category

indep3 <- list("relevel(imm_minorite, 2)")

indep4 <- list("relevel(imm_minorite, 2) + relevel(sexe, 2) + length_stay_2 + gpage + niv_etude + statut_mat + statut_emploi") 
indep4_bis <- list("relevel(imm_minorite, 2) + relevel(sexe, 2) + length_stay_2 + gpage + niv_etude + statut_mat")

```

## Different models

- Model with dependent vars 1

model1.1 : only one independent variable (imm_minorite), reference : majority
model1.2 : control all independent variable (imm_minorite), reference : majority
model1.3 : only one independent variable (imm_minorite), reference : immigrant not visible
model1.4 : control all independent variable (imm_minorite), reference : immigrant not visible

- Modele unique

model_u1.1 : only one independent variable (imm_minorite), reference : majority
model_u1.2 : control all independent variable (imm_minorite), reference : majority
model_u1.3 : only one independent variable (imm_minorite), reference : immigrant not visible
model_u1.4 : control all independent variable (imm_minorite), reference : immigrant not visible

model1.2 : only one independent variable (imm_minorite), reference : immigrant not visible

## Distribution of the dependent variables

```{r}

freq(trajipaq_covid$f_role_parent)

```

## Model for all dependent variables except income and entente dans la famille

```{r}

model <- function(dep, indep){
  Map(function(x,y) glm(as.formula(paste(x,paste(list(y),collapse="+"))),data=trajipaq_covid, family="binomial"), dep, indep)
}

expo <- list("e_hors_dom~", "e_four_soin~", "e_quarantaine~", "e_expose~",
            "e_symptome~", "e_hospitalise~", "e_intensif~", "e_decede~")

ressources <- list("f_nourriture~", "f_medicaments~", "f_sante~", "f_assurance~",
                   "f_essentiel~")

ability_b <- list("f_separe~", "f_evenement~")

ability_o <- list("f_role_parent~", "f_entende~", "f_enfant_maladie~", "f_autre_enfant~", "f_pers_agees~")

#f_role_parent, f_entende, f_enfant_maladie, f_autre_enfant, f_pers_agees)

#cons <- list("f_separe~", "f_emmenage~", "f_quitte_dom~", "f_nourriture~", "f_medicaments~", "f_sante~", "f_assurance~", "f_evenement~", "f_essentiel~")

empl <- list("f_reduire~", "f_cesser~", "f_perdu~", "f_revenu~")

aide <- list("patt~", "pcu~")

independent <- list(indep1, indep2, indep3, indep4)
dependent <- list(expo, ressources, ability_b, aide)

model_global <- NULL
model_global <- list()

for (i in 1:length(dependent)) {
  for (j in 1:length(independent)) {
    model_i_j <- model(dependent[[i]], independent[[j]])
    model_global[[paste0("Model", i, j)]] <- model_i_j 
  }
}

cov_lab1 <- c("Non-racialised immigrant", "Racialised immigrant", "Female", "More than 5 years", "Born here", "35-44",
               "45-60", "Collegial or professional training", "University", "Maried", "Living together",
               "Separed, Divorced, Widow", "Unemployed", "Others")

cov_lab2 <- c("Majority group", "Racialised immigrant", "Female", "More than 5 years", "Born here", "35-44",
               "45-60", "Collegial or professional training", "University", "Maried", "Living together",
               "Separed, Divorced, Widow", "Unemployed", "Others")

#More than 5 years
# Result for exposition

dep_lab1 <- c("Work outside", "Health provider", "Quarantine", "Exposed", "Symptoms", "Hospitalised", "Intensive care", "Has dead")

stargazer(model_global[4], ci = TRUE, align = TRUE, type = "text", 
          dep.var.labels = c("Exposition"), column.labels = dep_lab1)

stargazer(model_global[1], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab1,
          dep.var.labels = c("Exposition"), column.labels = dep_lab1)

# stargazer(model_global[1], ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text", covariate.labels = cov_lab, dep.var.labels = c("Exposition"), column.labels = dep_lab1)

stargazer(model_global[2], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab1,
          dep.var.labels = c("Exposition"), column.labels = dep_lab1)

stargazer(model_global[3], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab2,
          dep.var.labels = c("Exposition"), column.labels = dep_lab1)

stargazer(model_global[4], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab2,
          dep.var.labels = c("Exposition"), column.labels = dep_lab1)

# On getting ressource

dep_lab2 <- c("Food", "Medicine", "Health care", "Family insurrance", "Other essentials")

stargazer(model_global[5], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab1,
          dep.var.labels = c("Difficulty getting..."), column.labels = dep_lab2)

stargazer(model_global[6], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab1,
          dep.var.labels = c("Difficulty getting..."), column.labels = dep_lab2)

stargazer(model_global[7], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab2,
          dep.var.labels = c("Difficulty getting..."), column.labels = dep_lab2)

stargazer(model_global[8], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab2,
          dep.var.labels = c("Difficulty getting..."), column.labels = dep_lab2)

# On ability to perform usual duties

dep_lab3 <- c("Separated", "Important event")

stargazer(model_global[9], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab1,
          dep.var.labels = c("Performing usual duties"), column.labels = dep_lab3)

stargazer(model_global[10], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab1,
          dep.var.labels = c("Performing usual duties"), column.labels = dep_lab3)

stargazer(model_global[11], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab2,
          dep.var.labels = c("Performing usual duties"), column.labels = dep_lab3)

stargazer(model_global[12], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab2,
          dep.var.labels = c("Performing usual duties"), column.labels = dep_lab3)

# Model for government assistance

dep_lab4 <- c("PATT", "PCU")

stargazer(model_global[13], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab1,
          dep.var.labels = c("Government assistance"), column.labels = dep_lab4)

stargazer(model_global[14], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab1,
          dep.var.labels = c("Government assistance"), column.labels = dep_lab4)

stargazer(model_global[15], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab2,
          dep.var.labels = c("Government assistance"), column.labels = dep_lab4)

stargazer(model_global[16], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab2,
          dep.var.labels = c("Government assistance"), column.labels = dep_lab4)

```

## Model for assistance with appropriate data

```{r}

freq(trajipaq_covid$f_revenu)

trajipaq_covid_aide <- trajipaq_covid %>% 
  filter(f_reduire == "oui" | f_cesser == "oui" | f_perdu == "oui" | f_revenu == "oui")

# 723 individus remplissent cette condition

ctable(trajipaq_covid_aide$imm_minorite, trajipaq_covid_aide$patt)

independent <- list(indep1, indep2, indep3, indep4)
aide <- list("patt~", "pcu~")

# New dependent variable

ctable(trajipaq_covid_aide$pcu, trajipaq_covid_aide$patt)

trajipaq_covid_aide <- trajipaq_covid_aide %>% 
  mutate(assistance = case_when(
    patt == "Oui" | pcu == "Oui" ~ 1,
    patt == "Non" & pcu == "Non" ~ 0,
    TRUE ~ NA_real_

  ))

freq(trajipaq_covid_aide$assistance)

# Model

aide1 <- list("assistance~")

model <- function(dep, indep){
  Map(function(x,y) glm(as.formula(paste(x,paste(list(y),collapse="+"))),data=trajipaq_covid_aide, family="binomial"), dep, indep)
}

model_aide <- NULL
model_aide <- list()

for (i in 1:length(independent)) {
  model_i <- model(aide1, independent[[i]])
  model_aide[[paste0("Model", i)]] <- model_i 
}

cov_lab1 <- c("Non-racialised immigrant", "Racialised immigrant", "Female", "More than 5 years", "Born here", "35-44",
               "45-60", "Collegial or professional training", "University", "Maried", "Living together",
               "Separed, Divorced, Widow", "Unemployed", "Others")

cov_lab2 <- c("Majority group", "Racialised immigrant", "Female", "More than 5 years", "Born here", "35-44",
               "45-60", "Collegial or professional training", "University", "Maried", "Living together",
               "Separed, Divorced, Widow", "Unemployed", "Others")

#dep_lab4 <- c("PATT", "PCU")
dep_lab4 <- c("Assistance received")

stargazer(model_aide[1], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab1,
          dep.var.labels = c("Government assistance"), column.labels = dep_lab4)

stargazer(model_aide[2], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab1,
          dep.var.labels = c("Government assistance"), column.labels = dep_lab4)

stargazer(model_aide[3], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab2,
          dep.var.labels = c("Government assistance"), column.labels = dep_lab4)

stargazer(model_aide[4], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab2,
          dep.var.labels = c("Government assistance"), column.labels = dep_lab4)

```

## model for emploi

### Emploi

```{r}

independent_bis <- list(indep1, indep2_bis, indep3, indep4_bis)

dep1 <- list("f_reduire~", "f_cesser~", "f_perdu~", "f_revenu~")

model_emp <- NULL
model_emp <- list()

for (i in 1:length(independent)) {
  model_i <- model(dep1, independent_bis[[i]])
  model_emp[[paste0("Model", i)]] <- model_i 
}

cov_lab1 <- c("Non-racialised immigrant", "Racialised immigrant", "Female", "More than 5 years", "Born here", "35-44",
               "45-60", "Collegial or professional training", "University", "Maried", "Living together",
               "Separed, Divorced, Widow")

cov_lab2 <- c("Majority group", "Racialised immigrant", "Female", "More than 5 years", "Born here", "35-44",
               "45-60", "Collegial or professional training", "University", "Maried", "Living together",
               "Separed, Divorced, Widow")

#, "Unemployed", "Others"
dep_lab <- c("Limit", "Stop", "Lost", "Income reduced")

stargazer(model_emp[1], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab1,
          dep.var.labels = c("Consequences on employment and income"), column.labels = dep_lab)

stargazer(model_emp[2], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab1,
          dep.var.labels = c("Consequences on employment and income"), column.labels = dep_lab)

stargazer(model_emp[3], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab2,
          dep.var.labels = c("Consequences on employment and income"), column.labels = dep_lab)

stargazer(model_emp[4], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab2,
          dep.var.labels = c("Consequences on employment and income"), column.labels = dep_lab)

```

### Entente dans la famille - on the ability to perform duties

```{r}

library(MASS)

dep1 <- list("f_role_parent~", "f_entende~", "f_enfant_maladie~", "f_autre_enfant~", "f_pers_agees~")

model <- function(dep, indep){
  Map(function(x,y) polr(as.formula(paste(x,paste(list(y),collapse="+"))), method = c("logistic"), Hess = TRUE, data=trajipaq_covid), dep, indep) }

independent <- list(indep1, indep2, indep3, indep4)

model_ent <- NULL
model_ent <- list()

for (i in 1:length(independent)) {
  model_i <- model(dep1, independent[[i]])
  model_ent[[paste0("Model", i)]] <- model_i 
}

#cov_lab1 <- c("Immigrant and a visible minority", "Immigrant not a visible minority", "Others", "Female", "35-44",
#               "45-60", "Collegial or professional training", "University", "Maried", "Living together",
#               "Separed, Divorced, Widow", "Unemployed", "Others")

dep_lab <- c("Parenting", "Get along", "Care for illness child", "Care for other child", "Care for elderly or disabled")

cov_lab1 <- c("Non-racialised immigrant", "Racialised immigrant", "Female", "More than 5 years", "Born here", "35-44",
               "45-60", "Collegial or professional training", "University", "Maried", "Living together",
               "Separed, Divorced, Widow", "Unemployed", "Others")

cov_lab2 <- c("Majority group", "Racialised immigrant", "Female", "More than 5 years", "Born here", "35-44",
               "45-60", "Collegial or professional training", "University", "Maried", "Living together",
               "Separed, Divorced, Widow", "Unemployed", "Others")

stargazer(model_ent[1], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab1,
          dep.var.labels = c("Ability to ..."), column.labels = dep_lab)

stargazer(model_ent[2], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab1,
          dep.var.labels = c("Ability to ..."), column.labels = dep_lab)

stargazer(model_ent[3], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab2,
          dep.var.labels = c("Ability to ..."), column.labels = dep_lab)

stargazer(model_ent[4], ci = TRUE, align = TRUE, type = "text", covariate.labels = cov_lab2,
          dep.var.labels = c("Ability to ..."), column.labels = dep_lab)

```

<!--

## Interaction model

```{r}

model <- function(dep, indep){
  Map(function(x,y) glm(as.formula(paste(x,paste(list(y),collapse="+"))),data=trajipaq_covid, family="binomial"), dep, indep)
}

independent <- list(indep1, indep2, indep3, indep4)

model_expo <- NULL
model_expo <- list()

for (i in 1:length(independent)) {
  model_i <- model(dep1, independent[[i]])
  model_expo[[paste0("Model", i)]] <- model_i 
}

cov_lab <- c("Immigrant and a visible minority", "Immigrant not a visible minority", "Others", "Female", "35-44",
               "45-60", "Collegial or professional training", "University", "Maried", "Living together",
               "Separed, Divorced, Widow", "Unemployed", "Others")

stargazer(model_expo[1], ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text",
          covariate.labels = cov_lab)
stargazer(model_expo[2], ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text",
          covariate.labels = cov_lab)
stargazer(model_expo[3], ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text")
stargazer(model_expo[4], ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text")

# Modele unique

cov_lab_u <- c("Immigrant and a visible minority", "Immigrant not a visible minority", "Others", "Female", "35-44",
               "45-60", "Collegial or professional training", "University", "Maried", "Living together",
               "Separed, Divorced, Widow", "Unemployed", "Others", "Immigrant and a visible minority", "Majority group",
               "Others")

model_expo_u <- NULL
model_expo_u <- list()

for (i in 1:4) {
  model_i <- model(dep2, independent[[i]])
  model_expo_u[[paste0("Model", i)]] <- model_i 
}

stargazer(model_expo_u[1:4], apply.coef = exp, ci = TRUE, apply.ci = exp, align = TRUE, type = "text",
          covariate.labels = cov_lab_u)

```

### Consequences 1

```{r}

consequences <-
  trajipaq_covid %>% 
  dplyr::select(f_separe, f_emmenage, f_quitte_dom, f_nourriture, f_medicaments, f_sante, f_revenu, f_assurance, f_evenement, f_essentiel)

dfSummary(consequences, 
          plain.ascii  = FALSE, 
          style        = "grid", 
          graph.magnif = 0.75, 
          valid.col    = FALSE,
          tmp.img.dir  = "/tmp")

dep1 <- list("f_separe~", "f_emmenage~", "f_quitte_dom~", "f_nourriture~", "f_medicaments~", "f_sante~", "f_revenu~", "f_assurance~", "f_evenement~", "f_essentiel~")

model_cons <- NULL
model_cons <- list()

for (i in 1:length(independent)) {
  model_i <- model(dep1, independent[[i]])
  model_cons[[paste0("Model", i)]] <- model_i 
}

cov_lab <- c("Immigrant and a visible minority", "Immigrant not a visible minority", "Others", "Female", "35-44",
               "45-60", "Collegial or professional training", "University", "Maried", "Living together",
               "Separed, Divorced, Widow", "Unemployed", "Others")

stargazer(model_cons[1], ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text",
          covariate.labels = cov_lab)
stargazer(model_cons[2], ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text",
          covariate.labels = cov_lab)
stargazer(model_cons[3], ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text")
stargazer(model_cons[4], ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text")

#model3 <- Map(function(x,y) glm(as.formula(paste(x,paste(list(y),collapse="+"))),data=trajipaq_covid, family="binomial"), dep2, indep1)

#stargazer(model3, ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text")

#model4 <- Map(function(x, y) glm(as.formula(paste(x,paste(list(y),collapse="+"))),data=trajipaq_covid, family="binomial"), dep2, indep2) 
#stargazer(model4, ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text")

```

### Emploi

```{r}

emploi <-
  trajipaq_covid %>% 
  dplyr::select(f_reduire, f_cesser, f_perdu)

dfSummary(emploi, 
          plain.ascii  = FALSE, 
          style        = "grid", 
          graph.magnif = 0.75, 
          valid.col    = FALSE,
          tmp.img.dir  = "/tmp")

dep1 <- list("f_reduire~", "f_cesser~", "f_perdu~")

model_emp <- NULL
model_emp <- list()

for (i in 1:length(independent)) {
  model_i <- model(dep1, independent[[i]])
  model_emp[[paste0("Model", i)]] <- model_i 
}

cov_lab <- c("Immigrant and a visible minority", "Immigrant not a visible minority", "Others", "Female", "35-44",
               "45-60", "Collegial or professional training", "University", "Maried", "Living together",
               "Separed, Divorced, Widow", "Unemployed", "Others")
dep_lab <- c("Limit", "Stop", "Lost")

stargazer(model_emp[1], ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text",
          covariate.labels = cov_lab, dep.var.labels = c("Consequences on employment"), column.labels = dep_lab)
stargazer(model_emp[2], ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text",
          covariate.labels = cov_lab)
stargazer(model_emp[3], ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text")
stargazer(model_emp[4], ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text")

model5 <- Map(function(x,y) glm(as.formula(paste(x,paste(list(y),collapse="+"))),data=trajipaq_covid, family="binomial"), dep3, indep1) 
stargazer(model5, ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text")

model6 <- Map(function(x, y) glm(as.formula(paste(x,paste(list(y),collapse="+"))),data=trajipaq_covid, family="binomial"), dep3, indep2) 
stargazer(model6, ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text")

```

### Bien-être

```{r}

bienetre <-
  trajipaq_covid %>% 
  dplyr::select(be_exercice, be_manger, be_dormir, be_anxiete, be_humeur)

dfSummary(bienetre, 
          plain.ascii  = FALSE, 
          style        = "grid", 
          graph.magnif = 0.75, 
          valid.col    = FALSE,
          tmp.img.dir  = "/tmp")

freq(trajipaq_covid$be_exercice)

dep5 <- list("be_exercice~", "be_manger~", "be_dormir~", "be_anxiete~", "be_humeur~")

model9 <- Map(function(x,y) polr(as.formula(paste(x,paste(list(y),collapse="+"))), method = c("logistic"), Hess = TRUE, data=trajipaq_covid), dep5, indep1) 
stargazer(model9, ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text")

model10 <- Map(function(x, y) polr(as.formula(paste(x,paste(list(y),collapse="+"))), method = c("logistic"), Hess = TRUE, data=trajipaq_covid), dep5, indep2) 
stargazer(model10, ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text")

```

### Aide reçue

```{r}

aide <-
  trajipaq_covid %>% 
  dplyr::select(patt, pcu)

dfSummary(aide, 
          plain.ascii  = FALSE, 
          style        = "grid", 
          graph.magnif = 0.75, 
          valid.col    = FALSE,
          tmp.img.dir  = "/tmp")

dep6 <- list("patt~", "pcu~")

model11 <- Map(function(x,y) glm(as.formula(paste(x,paste(list(y),collapse="+"))),data=trajipaq_covid, family="binomial"), dep6, indep1) 
stargazer(model11, ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text")

model12 <- Map(function(x, y) glm(as.formula(paste(x,paste(list(y),collapse="+"))),data=trajipaq_covid, family="binomial"), dep6, indep2) 
stargazer(model12, ci = TRUE, apply.coef = exp, apply.ci = exp, align = TRUE, type = "text")

```

-->